---
title: "CSDS341_Project_accessingMySQL"
author: "Luke Zhang (rxz330)"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  rmdformats::html_clean:
    highlight: kate
    toc: yes
    toc_float: yes
    number_sections: yes
    thumbnails: no
    fig_width: 8
---
# Preliminaries 

## Load Packages 
The connection between `R` and `MySQL` will require the following packages.
```{r packages, message=FALSE, warning=FALSE}
library(RMySQL)
library(RODBC)
library(knitr)
library(tidyverse)
```

## Establish Connection to mySQL
Before starting query the database, we need to set up a link about our connection. We stored our connection in variable `sql_link`. 
```{r connectio_to_sql, message=FALSE}
# Establish Connection to MySQL
sql_link <- dbConnect(RMySQL::MySQL(),
                      dbname='airline',
                      host='localhost', 
                      port=3306,
                      user='root', 
                      password='Zjm200249@szboy')

# List All the Tables in the Schema
dbListTables(sql_link)
```

# Queries & SQL

## Easy Queries & SQL
Query 01: Find the first name(`first_name`), last name(`last_name`), date of birth(`dob`), and the credits(`credits`) of all male travelers who are U.S. citizen. 
```{r}
sql_01 <- "SELECT first_name, last_name, dob, credits
           FROM   travelers
           WHERE  citizenship = 'US' 
                  AND 
                  gender = 'M'
                  AND 
                  dob < '1985-01-01'"

result_01 <- dbSendQuery(sql_link, sql_01)

output_01 <- fetch(result_01)
output_01 %>% kable()
```

Query 02: Find the passport number(`passport_no`) and first name(`first_name`) of all Chinese Traveler who has already booked at least one Plane Ticket. The result table should also include `citizenship` as a check.
```{r}
sql_02 <-	"SELECT passport_no, first_name, citizenship
		       FROM   Travelers
		       WHERE  (Citizenship = 'CN') 
		              AND 
		              (user_id IN (SELECT    traveler_id
		                           FROM      Tickets_book_for
		                           GROUP BY  traveler_id
		                           HAVING    count(ticket_id) >= 1))"

result_02 <- dbSendQuery(sql_link, sql_02)

output_02 <- fetch(result_02)
output_02 %>% kable()
```

Query 02: Find the flight number(`flight_no`), departure airport(`dept_iata_code`), arrival airport(`ariv_iata_code`) of all flights that operated by China Eastern Airlines and has a capacity of 30 seats in first class cabin. 
	% query starts here
```{r}
sql_03 <- "SELECT flight_no, dept_iata_code, ariv_iata_code
           FROM   Flights_ariv_dept NATURAL JOIN Aeroplanes_belong
           WHERE  (Aeroplanes_belong.fir_capacity = 30) 
           AND    (regis_no IN (
              SELECT regis_no
              FROM   Aeroplanes_belong NATURAL JOIN Airlines
              WHERE  Airlines.company_name = 'China Eastern Airlines'))"

result_03 <- dbSendQuery(sql_link, sql_03)

output_03 <- fetch(result_03)
output_03 %>% kable()
```
