---
title: "Database"
author: 'Quynh Nguyen'
email: --
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library("readxl")
library("stringr")
library("randomNames")
library("stringr")
library(readr)
library(dplyr)
library(tibble)
library("data.table")
library(lubridate)
library("xlsx")
library(nycflights13)
```

```{r 1. TRAVELERS}
country <- data.frame(read_excel("Downloads/country_code.xlsx"))
genders <- c("F", "M", "U")
df <- data.frame(GENDER=sample(genders, 50000, replace=TRUE))
names <- randomNames(gender=df$GENDER, name.sep = " ")
travelers <- data.frame(str_split_fixed(names, " ", 2))
colnames(travelers) <- c("first_name", "last_name")
travelers$middle_name<- randomNames(gender=df$GENDER, name.sep = " ", which.names = "first")
travelers$gender <- df$GENDER
travelers$dob <- sample(seq(as.Date('1903/01/01'), as.Date('2022/02/01'), by="day"), nrow(travelers), replace = TRUE)
travelers$credits <- sample(0:6000, size = 50000, replace = TRUE)

myFun <- function(n = 100000) {
  a <- do.call(paste0, replicate(1, sample(LETTERS, n, TRUE), FALSE))
  paste0(a, sprintf("%04d", sample(99999999, n, TRUE)))
}

travelers$user_id <- sample(100000:9999999999,nrow(travelers), replace = FALSE)
travelers$passport_no <- myFun(nrow(travelers))

travelers$citizenship <- sample(country$Alpha.2.code,nrow(travelers), replace = TRUE)
travelers <- travelers[,c("user_id","first_name", "middle_name", "gender", "dob", "credits", "passport_no", "citizenship" )]

head(travelers)

write.xlsx(travelers, file = 'travelers.xlsx', row.names = FALSE)
```

```{r 2. CREW}
genders <- c("F", "M", "U")
df <- data.frame(GENDER=sample(genders, 5000, replace=TRUE))
names <- randomNames(gender=df$GENDER, name.sep = " ")
crew <- data.frame(str_split_fixed(names, " ", 2))
colnames(crew) <- c("first_name", "last_name")
crew$middle_name<- randomNames(gender=df$GENDER, name.sep = " ", which.names = "first")
crew$gender <- df$GENDER
crew$dob <- sample(seq(as.Date('1962/01/01'), as.Date('1999/12/31'), by="day"), nrow(crew), replace = TRUE)
crew$ssn <- sample(10000000:99999999, size = nrow(crew), replace = FALSE)
crew$user_id <- sample(0:99999, nrow(crew), replace = FALSE) 
crew$total_distance <- sample(0:430000, nrow(crew), replace = TRUE)
crew['salary'] <-NA
for (i in 1: nrow(crew)) {
  if (crew[i,]$total_distance <= 1000) {
    crew[i,]$salary <- 100000
  }
  if(crew[i,]$total_distance > 1000 & crew[i,]$total_distance <= 50000){
    crew[i,]$salary <- 150000
  }
  if(crew[i,]$total_distance > 50000 & crew[i,]$total_distance <= 100000){
    crew[i,]$salary <- 180000
  }
  else{
    crew[i,]$salary <- 200000
  }
}

crew <- crew[,c("user_id", "first_name", "middle_name", "last_name", "gender", "dob", "ssn", "salary", "total_distance")]

write.xlsx(crew, file = 'update_crew.xlsx', row.names = FALSE)
```

```{r 3. US Airports}
raw_airport <- read.csv("https://davidmegginson.github.io/ourairports-data/airports.csv")

#keep large airport 
ra <- raw_airport[which(raw_airport$type == 'medium_airport' | raw_airport$type == 'large_airport'),]

##only keep airport that have IATA code
ra <- ra[!(ra$iata_code == ""),]

#create data frame 
airport <- data.frame(ra$iata_code, ra$name, ra$iso_country)
colnames(airport) <- c("iata_code", "airport_name", "country")

##Remove airport that has no country
airport <- airport[!is.na(airport$country), ]

##Remove army air field, air force, and air station
airport <- airport %>% filter(!grepl('Army', airport_name))
airport <- airport %>% filter(!grepl('Station', airport_name))
airport <- airport %>% filter(!grepl('Air Force', airport_name))
airport <- airport %>% filter(!grepl('Airfield', airport_name))
airport <- airport %>% filter(!grepl('Carr Field', airport_name))

#random assign weather for each airport
weather <- c('Sunny' , 'Mostly Sunny', 'Partly Cloudy', 'Cloudy', 'Rainy', 'Heavy Rainy', 'Foggy', 'Snowy', 'Heavy Snowy', 'Frost')

weather <- data.frame(weather)

airport$weather <- sample(weather$weather, nrow(airport), replace = TRUE)

#random assign status of the airport
status <- c('Free' , 'Normal', 'Busy', 'Small-Scale Delay', 'Large-Scale Delay')
status <- data.frame(status)
airport$airport_status <- sample(status$status, nrow(airport), replace = TRUE)

##Only get US International Airport
us_airport <- airport[airport$country == 'US', ]
a1 <- us_airport %>% filter(grepl('International', airport_name))
a2 <- us_airport %>% filter(grepl('Alliance', airport_name))

us_intl_airport <- data.frame(rbind(a1, a2))
write.xlsx(us_airport, file = 'us_airport.xlsx', row.names = FALSE)
write.xlsx(us_intl_airport, file = 'us_intl_airport.xlsx', row.names = FALSE)
```

```{r 4. US Airlines}
##US Airlines only
raw_us_airline <- read_excel("Downloads/us_airline.xlsx")
raw_us_airline <- data.frame(na.omit(raw_us_airline))
colnames(raw_us_airline) <- c("company_name", "IATA", "ICAO")

cape_air <- c("Cape Air", "9K", "KAP")
republic_air <- c("Republic Airline", "YX", "RPA")
raw_us_airline <- rbind(raw_us_airline, cape_air)
raw_us_airline <- rbind(raw_us_airline, republic_air)

##Conver letter to number
LETTER2num <- function(x) {utf8ToInt(x) - utf8ToInt("A") + 1L}
icao <- raw_us_airline[,3]
id <- c()
a <- c()

##convert ICAO of each airlines into its id
for (i in 1: length(icao)){
  a <- icao[i]
  a <- strsplit(a, split ="")
  a <- unlist(a)
  b1 <- LETTER2num(a[1])
  b2 <- LETTER2num(a[2])
  b3 <- LETTER2num(a[3])
  id[i] <- strtoi(paste(b1, b2, b3, sep =""))
}

raw_us_airline$company_id <- id


us_airline <- raw_us_airline[, c("company_id", "company_name")] ##export this to excel file later
write.xlsx(us_airline, file = 'us_airline.xlsx', row.names = FALSE)
```

```{5. Schedules}
dat_airport <- data.frame(read_csv("Downloads/Airports2.csv", show_col_types = FALSE))
dat_airport <- dat_airport[!(dat_airport$Distance == 0),]

dis_between <- unique(dat_airport[c("Origin_airport", "Destination_airport", "Distance")])
dis_between <- dis_between[dis_between$Origin_airport %in% us_airport$iata_code,]
dis_between <- dis_between[dis_between$Destination_airport%in% us_airport$iata_code,]

dt <-  as.data.table(dis_between)
dis_between <- data.frame(dt[, mean(Distance), by = list(Origin_airport, Destination_airport)])
colnames(dis_between) <- c("Origin Airport", "Destination Airport", "Distance")

##calculation between two airports
dis_between$duration <- round((dis_between$Distance*18)/60)

d
##create flights
is_between$freq <- sample(1:3, nrow(dis_between), replace = TRUE)
flight  <- data.frame('Origin Airport' = character(), 
                      'Destination Airport' = character(), 
                       'Distance' = integer(),
                        stringsAsFactors=FALSE)
for (i in 1: nrow(dis_between)){
  row <- i
  time <- dis_between[i,]$freq
  data <- dis_between[rep(row, time),]
  flight <- rbind(flight, data)
}


##Create depature date + time

possible_dates <- seq(as.POSIXct('2019/01/01'), as.POSIXct('2022/05/31'), by="15 mins")
possible_dates <- possible_dates[hour(possible_dates) < 21 & hour(possible_dates) >= 9]
depature <- sample(possible_dates, nrow(flight), replace = TRUE)
dept_time<- data.frame(str_split_fixed(depature, " ", 2))

flight <- cbind(flight, dept_time)
colnames(flight) <- c('Origin Airport', 'Destination Airport', 'Distance', 'Duration', 'Frequency', 'Depature Day', 'Depature Time')

flight$flight_id <- sample(1000:999999, nrow(flight), replace = FALSE)

##create schedule table - but still lack of arrival time and arrival day
schedule <- flight[,c("flight_id", 'Depature Day', 'Depature Time')]
write.xlsx(schedule, file = 'schedule.xlsx', row.names = FALSE)
```

```{r 6. US Aeroplanes_belong}
```{r 6. US Aeroplanes_belong}
us_fleet_raw <-  data.frame(read_excel("Downloads/us_fleet.xlsx"))
colnames(us_fleet_raw) <- c("Customer.Name", "Model.Series", "Order.Total")
us_fleet <- data.frame('Customer' = character(), 
                             'Model.Series' = character(), 
                             'Order.Total' = integer(),
                             stringsAsFactors=FALSE)
for (i in 1: nrow(us_fleet_raw)){
  row <- i
  time <- us_fleet_raw[i,]$Order.Total
  data <- us_fleet_raw[rep(row, time),]
  us_fleet <- rbind(us_fleet, data)
}

us_fleet <- us_fleet[, c("Customer.Name", "Model.Series")]

us_fleet['fir_capacity'] <- NA
us_fleet['bus_capacity'] <- NA
us_fleet['eco_capacity'] <- NA

##assign capacity for each type of fleets
for (i in 1: nrow(us_fleet)){
  if(str_detect(us_fleet[i,]$Model.Series,'Boeing 737')){
     us_fleet[i,]$fir_capacity <- 16
     us_fleet[i,]$bus_capacity <- 24
    us_fleet[i,]$eco_capacity <- 138
  }
  
  if(str_detect(us_fleet[i,]$Model.Series,'Boeing 757')){
    us_fleet[i,]$fir_capacity <- 16
    us_fleet[i,]$bus_capacity <- 45
    us_fleet[i,]$eco_capacity <- 108
  }
  if(str_detect(us_fleet[i,]$Model.Series,'Boeing 767')){
      us_fleet[i,]$fir_capacity <- 30
      us_fleet[i,]$bus_capacity <- 49
      us_fleet[i,]$eco_capacity <- 135
  }
  if(str_detect(us_fleet[i,]$Model.Series,'Boeing 777')){
      us_fleet[i,]$fir_capacity <- 60
      us_fleet[i,]$bus_capacity <- 24
      us_fleet[i,]$eco_capacity <- 266
  }
  if(str_detect(us_fleet[i,]$Model.Series,'Boeing 787')){
      us_fleet[i,]$fir_capacity <- 36
      us_fleet[i,]$bus_capacity <- 70
      us_fleet[i,]$eco_capacity <- 113
  }
  
  if(str_detect(us_fleet[i,]$Model.Series,'Embraer 195')){
      us_fleet[i,]$fir_capacity <- 0
      us_fleet[i,]$bus_capacity <- 18
      us_fleet[i,]$eco_capacity <- 100
  }
  
   if(str_detect(us_fleet[i,]$Model.Series,'Embraer')){
      us_fleet[i,]$fir_capacity <- 0
      us_fleet[i,]$bus_capacity <- 18
      us_fleet[i,]$eco_capacity <- 100
   }
  
  if(str_detect(us_fleet[i,]$Model.Series,'Airbus A320')){
      us_fleet[i,]$fir_capacity <- 12
      us_fleet[i,]$bus_capacity <- 42
      us_fleet[i,]$eco_capacity <- 96
  }
  
  if(str_detect(us_fleet[i,]$Model.Series,'Airbus A321')){
      us_fleet[i,]$fir_capacity <- 16
      us_fleet[i,]$bus_capacity <- 15
      us_fleet[i,]$eco_capacity <- 156
  }
  if(str_detect(us_fleet[i,]$Model.Series,'Airbus A319')){
      us_fleet[i,]$fir_capacity <- 12
      us_fleet[i,]$bus_capacity <- 36
      us_fleet[i,]$eco_capacity <- 78
   }
  
  if(str_detect(us_fleet[i,]$Model.Series,'Airbus A220')){
      us_fleet[i,]$fir_capacity <- 12
      us_fleet[i,]$bus_capacity <- 15
      us_fleet[i,]$eco_capacity <- 82
  }
  
  if(str_detect(us_fleet[i,]$Model.Series,'Airbus A330')){
      us_fleet[i,]$fir_capacity <- 34
      us_fleet[i,]$bus_capacity <- 40
      us_fleet[i,]$eco_capacity <- 219
  }
  
  if(str_detect(us_fleet[i,]$Model.Series,'Airbus A350')){
      us_fleet[i,]$fir_capacity <- 48
      us_fleet[i,]$bus_capacity <- 21
      us_fleet[i,]$eco_capacity <- 224
  }
  if(str_detect(us_fleet[i,]$Model.Series,'Boeing 717')){
      us_fleet[i,]$fir_capacity <- 12
      us_fleet[i,]$bus_capacity <- 20
      us_fleet[i,]$eco_capacity <- 78
   }
  
  if(str_detect(us_fleet[i,]$Model.Series,'De Havilland Canada Dash 8-400')) {
    us_fleet[i,]$fir_capacity <- 0
    us_fleet[i,]$bus_capacity <- 7
    us_fleet[i,]$eco_capacity <- 60
   }
}

##match id with its airline
id <- c()
name <- NA
for (a in 1:nrow(us_fleet)) {
  for (b in 1:nrow(us_airline)){
    if(us_fleet[a,]$Customer.Name == us_airline[b,]$company_name){
      name <- us_airline[b,]$company_name
      id[a] <- us_airline$company_id[ us_airline$company_name == name]
    }
  }
}

us_fleet$company_id <- id

##generate random regis no number
myFun <- function(n = 100000) {
  a <- do.call(paste0, replicate(1, sample(LETTERS, n, TRUE), FALSE))
  paste0(a, sprintf("%04d", sample(999999, n, TRUE)))
}

us_fleet$regis_no <- myFun(nrow(us_fleet))

##rename columns
colnames(us_fleet) <- c("Customer.Name", "aircratft_type", "fir_capacity", "bus_capacity", "eco_capacity", "company_id", "regis_no")

##reordercolumns
us_fleet <- us_fleet[,c("regis_no", "fir_capacity", "bus_capacity", "eco_capacity", "aircratft_type", "company_id")]


write.xlsx(us_fleet, file = 'us_aeroplanes.xlsx', row.names = FALSE)
```

```{r 7.Flights_ariv_dept}
status <- c("On-Time" , "Delay", "Cancel")
flight_no <- flight$flight_id

regis_no <- us_fleet$regis_no
dept_iata_code <- flight$`Origin Airport`
ariv_iata_code <- flight$`Destination Airport`

tickets <- data.frame(flight_no, dept_iata_code, ariv_iata_code)
tickets$regis_no <- sample(regis_no, nrow(tickets), replace = TRUE)
tickets$flight_status <- sample(status, nrow(tickets), replace = TRUE)

flights_ariv_dept <- tickets[,c("regis_no", "flight_no", "flight_status", "ariv_iata_code", "dept_iata_code")]

```