---
title: "data_retrieval_&_pre-processing"
author: "Luke Zhang (rxz330)"
date: "`r Sys.Date()`"
output: html_document
---

# Preliminaries
```{r}
library(knitr)
library(dplyr)
library(tidyverse)
```

# Airports Table Data 
This section retrieves iata code and corresponding name of airports in the United States. 
```{r, warning=FALSE}
install.packages("jsonlite", repos="https://cran.rstudio.com/")
library("jsonlite")

# get list of all resources
json_file <- 'https://datahub.io/core/airport-codes/datapackage.json'
json_data <- fromJSON(paste(readLines(json_file), collapse = "")) 

# print all tabular data(if exists any)
for(i in 1:length(json_data$resources$datahub$type)){
  if(json_data$resources$datahub$type[i]=='derived/csv'){
    path_to_file = json_data$resources$path[i]
    raw_iata_data <- read.csv(url(path_to_file))
    print(data)
  }
}
```

This section perform data pre-processing (i.e. selecting the airports which has the iata code). 
Use the code `iata_airport %>% group_by(type) %>% count(type)` we can find out that there are 6 different airports: `closed`($$n = 262$$), `heliport` ($$n = 82$$), `seaplane_base` ($$n = 147$$), `small_airport`($$n = 4249$$), `medium_airport`($$n = 3849$$), `large_airport`($$n = 602$$). We only want to include the large airports.

```{r, warning=FALSE}
iata_airport <- raw_iata_data %>% 
  filter(complete.cases(name,iata_code,iso_country,type)) %>%
  filter(iata_code != "" & nchar(iso_country) == 2 ) %>%
  filter(type =="large_airport") %>% 
  rename(airport_name = name, country = iso_country) %>% 
  select(airport_name, iata_code, country) 
```

The `country` label is not correct for some airports. The following code help to fix it. 
```{r, warning=FALSE}
iata_airport <- iata_airport %>% 
  filter(country == replace(country, country == 'HK', 'CN'))
iata_airport <- iata_airport %>% 
  filter(country == replace(country, country == 'TW', 'CN'))
iata_airport <- iata_airport %>% 
  filter(country == replace(country, country == 'MO', 'CN'))
```

`weather` can have 10 different values: `Sunny`, `Mostly Sunny`, `Partly Cloudy`, `Cloudy`, `Rainy`, `Heavy Rainy`, `Foggy`, `Snowy`, `Heavy Snowy`, `Frost`. Additionally, `airport_status` can have 5 different values: `Free`, `Normal`, `Busy`, `Small-Scale Delay`, `Large-Scale Delay`. 
```{r}
set.seed(2022341)
iata_num <- dim(iata_airport)[1]

# set up weather data - randomly simulated 
weather1 <- round(runif(floor(iata_num),1,4))
weather2 <- round(runif(ceiling(iata_num*1/10),5,10))
weather <- c(weather1, weather2)

weather <- as.data.frame(weather) %>%
  slice_sample(., prop = iata_num/length(weather))

# set up status data
airport_status <- round(runif(floor(iata_num),1,5))

# combine the weather and status with the original data
iata_airport_bind <- cbind(iata_airport, weather, airport_status)

iata_airport_rename <- iata_airport_bind %>% 
  mutate(weather = replace(weather, weather == '1', 'Sunny'), 
         weather = replace(weather, weather == '2', 'Mostly Sunny'),
         weather = replace(weather, weather == '3', 'Partly Cloudy'),
         weather = replace(weather, weather == '4', 'Cloudy'),
         weather = replace(weather, weather == '5', 'Rainy'),
         weather = replace(weather, weather == '6', 'Heavy Rainy'),
         weather = replace(weather, weather == '7', 'Foggy'),
         weather = replace(weather, weather == '8', 'Snowy'),
         weather = replace(weather, weather == '9', 'Heavy Snowy'),
         weather = replace(weather, weather == '10', 'Frost'),
         airport_status = replace(airport_status, airport_status == '1', 'Free'), 
         airport_status = replace(airport_status, airport_status == '2', 'Normal'),
         airport_status = replace(airport_status, airport_status == '3', 'Busy'),
         airport_status = replace(airport_status, airport_status == '4', 'Small-Scale Delay'),
         airport_status = replace(airport_status, airport_status == '5', 'Large-Scale Delay')
         )

final_airports_data <- iata_airport_rename %>% 
  select(iata_code, airport_name, country, weather, airport_status)
```

Export the `Airport` Data. 
```{r}
write.csv(final_airports_data, "C:/Users/luke_/OneDrive/Desktop/csds341-project-airlineDatabase/CSDS341_Project_Data_Retrieval/csv_data/airports.csv",row.names = FALSE)
```

## Airlines and Hub Table Data

Load the raw airlines table data.
```{r}
raw_airline_data <- read.csv("C:/Users/luke_/OneDrive/Desktop/csds341-project-airlineDatabase/CSDS341_Project_Data_Retrieval/raw_data/raw_airlines_data.csv", header = TRUE, sep =",")
```

The final airline table should have the following attributes: `company_id` and `company_name`.   
```{r}
airline <- raw_airline_data %>% 
  filter(Active == "Y") %>%
  rename(company_name = X.Name, country_name = Country) %>% 
  select(company_name, country_name)
```

Load the `raw_country_code` dataset for matching the airlines to their country.
```{r}
raw_country_code <- read.csv("C:/Users/luke_/OneDrive/Desktop/csds341-project-airlineDatabase/CSDS341_Project_Data_Retrieval/raw_data/raw_country_code.csv")

country_code <- raw_country_code %>%
  rename(country_name = ï..name) %>% 
  filter(complete.cases(country_name, code)) %>%
  filter(country_name != "" | code != "") %>%
  select(country_name, code)
```

Merge the `country_code` and `airline` table. 
```{r}
airline_country <- merge(airline, country_code, by = "country_name", all = FALSE) %>% 
  select(company_name,code) %>% 
  filter(code != "CN")

airline_country[nrow(airline_country)+1,] = c("China Eastern Airlines","CN")
airline_country[nrow(airline_country)+1,] = c("China Southern Airlines","CN")
airline_country[nrow(airline_country)+1,] = c("Air China","CN")
airline_country[nrow(airline_country)+1,] = c("Hainan Airlines","CN")
```

Create `id` for airline companies.
```{r}
set.seed(2022341)
company_id <- sample(101:999,nrow(airline_country), replace = FALSE)

airport_country_id <- cbind(company_id, airline_country)
```

Export the Airlines Data
```{r}
final_airlines_data <- airport_country_id %>% 
  select(company_id, company_name)

write.csv(final_airlines_data, "C:/Users/luke_/OneDrive/Desktop/csds341-project-airlineDatabase/CSDS341_Project_Data_Retrieval/csv_data/airlines.csv",row.names = FALSE)
```

Merge the `airport_country_id`(`code`) with `final_airports_data`(`country`). 
```{r}
airport_country_id_rename <- airport_country_id %>% 
  rename(country = code)

set.seed(2022341)
raw_hub <- merge(airport_country_id_rename, final_airports_data, by = "country") %>%
  slice_sample(., prop =3/4)

final_hub_data <- raw_hub %>%
  select(company_id, iata_code)
```

Export the hub data.
```{r}
write.csv(final_hub_data, "C:/Users/luke_/OneDrive/Desktop/csds341-project-airlineDatabase/CSDS341_Project_Data_Retrieval/csv_data/hub.csv",row.names = FALSE)
```


